---
title: "Extracting tidy samples from rethinking models"
author: "Matthew Kay"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    df_print: kable
vignette: >
  %\VignetteIndexEntry{Extracting tidy data from rstanarm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
<style type="text/css">
.kable-table table {
  margin-left: 0;
}
</style>


## Introduction
  
This vignette describes how to use the `tidybayes` package to extract tidy samples from `rethinking`. For a more general introduction to `tidybayes` and its use on more general-purpose sampling languages (like Stan and JAGS), see `vignette("tidybayes")`.
  
The default output data formats of popular samplers like JAGS and Stan often don't quite conform to the ideal of [tidy data](http://dx.doi.org/10.18637/jss.v059.i10). Output formats will often be in matrix form (requiring conversion for use with libraries like ggplot). `tidybayes` automates munging the samples into tidy formats.
  
  
## Setup
  
The following libraries are required to run this vignette:
  
```{r setup, message = FALSE, warning = FALSE}
library(magrittr)
library(dplyr)
library(tidybayes)
library(tidybayes.rethinking)
library(ggplot2)
library(ggstance)
library(rstan)
library(rethinking)
```

These options help Stan run faster:

```{r, eval=FALSE}
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

```{r hidden_options, include=FALSE}
# While the previous code chunk is the actual recommended approach,
# CRAN vignette building policy limits us to 2 cores, so we use at most
# 2 to build this vignette (but show the previous chunk to
# the reader as a best pratice example)
rstan_options(auto_write = TRUE)
options(mc.cores = min(2, parallel::detectCores()))

#ggplot options
theme_set(theme_light())

#figure options
knitr::opts_chunk$set(fig.width = 6, fig.height = 4)
```


## Example dataset

To demonstrate `tidybayes`, we will use a simple dataset with 10 observations from 5 conditions each:

```{r}
set.seed(5)
n = 10
n_condition = 5
ABC =
  data_frame(
    condition = rep(c("A","B","C","D","E"), n),
    response = rnorm(n * 5, c(0,1,2,1,-1), 0.5)
  )
```

A snapshot of the data looks like this:

```{r}
head(ABC, 10)
```
*(10 rows of `r nrow(ABC)`)*

This is a typical tidy format data frame: one observation per row. Graphically:

```{r}
ABC %>%
  ggplot(aes(y = condition, x = response)) +
  geom_point()
```

## Model

TBD.
